---
title: 'Analysis with Ensambles: competition data (transmittance values)'
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Use only 3 traits


* All columns are used
* The transmittance values are not transformed to absorbance values.
* Only have responses for the training set


## Compared to the older version: 

* ALl traits are scaled and centered
* PLS with optimal no of components
* Added lasso to stack regressions (enslasso)


```{r echo = FALSE, message=FALSE, warning= FALSE}
rm(list = ls())
set.seed(1979)
library(tidyverse)
library(dendextend)
library(glmnet) # for the ridge regression
library(pls)
library(ranger)
library(bartMachine)
library(e1071)
library(BKPC)
library(kernlab)
library(caret) # For ten fold crossvalidation
library(nnls)
library(effects)
library(lme4)
```


```{r echo = FALSE, message=FALSE, warning= FALSE, eval = FALSE}



tr <- read.csv("tr.csv")  


calcRMSE <- function(y, yhat){sqrt(mean((yhat - y)^2))}


ys <- tr[,1:3]
tr <- tr[, -c(1:3)]


a <- apply(tr,2,mean)
b <- apply(tr,2,sd)


tr <- scale(tr,a,b)

ys[,2] <- log(ys[,2])
# pairs(ys)
a <- apply(ys,2,mean, na.rm = TRUE)
b <- apply(ys,2,sd, na.rm = TRUE)

ys <- sweep(ys,2,a,"-")
ys <- sweep(ys,2,b,"/")

kfunc <-  rbfdot(sigma = 0.0005)




# pairs(log(ys)) # all non-normal
ys %>% pivot_longer(1:3, "key", "value") %>% ggplot(aes(x = value))+ 
  geom_histogram()+facet_wrap(vars(key), scales = "free")

# matplot(t(tr), type = "l", col = 9, lty = 1)
# matplot(t(te), type = "l", col = 9, lty = 1)


# pairs(tr[,1:10])



```


<!-- ## Look for clusters in the training data -->

```{r echo = FALSE, message=FALSE, warning= FALSE,  eval = FALSE}
D <- as.matrix(dist(tr,method="euclidean"))
# image(D)


h <- hclust(as.dist(D), method = "ward.D2")
d <- as.dendrogram(h)
# plot(d)

heatmap(D,
        Colv=as.dendrogram(h),     
        Rowv=as.dendrogram(h))


d2 <- color_branches(d,k=7, col=c(2,3,5,4, 6,7,8)) 
# auto-coloring 4 clusters of branches.
plot(d2)

# plot(reorder(d2, D, method = "OLO")) 

labl.Ave <- cutree(d2,7)
matplot(t(tr), type = "l", col = labl.Ave, lty = 1, ylab = "training data") # see the 7 clusters

# table(a$Farm[-id_te], labl.Ave)
# table(a$Breed[-id_te], labl.Ave)

#############################################################

```


<!-- ## Look for clusters in the training data -->

```{r echo = FALSE, message=FALSE, warning= FALSE,  eval = FALSE}
#############################################################

kfunc <-  rbfdot(sigma = 0.0005)
# kfunc <- laplacedot(sigma = 0.0001)
# kfunc <- anovadot(sigma = 1, degree = 1)
Ktrain <- kernelMatrix(kfunc, tr)
# image(Ktrain)





kpcTrain <- kPCA(Ktrain)
ys_tr <- ys

# plot the data projection on the principal components

# pairs(kpcTrain$KPCs[ , 1 : 6], col = labl.Ave)

newdat <- cbind(log(ys_tr), kpcTrain$KPCs[ , 1 : 6], labl.Ave)

newdat <- newdat %>% as.data.frame() %>% na.omit() %>% mutate(labl.Ave = as.factor(labl.Ave))
# pairs(newdat[,1:20],  col = newdat[,21])
# kpcTest <- predict(kpcTrain, Ktest)
# pairs(kpcTest[ , 1 : 6])
#############################################################

# correlation with responses

indexcor <- abs(cor(cbind(ys_tr,tr), use = "pairwise.complete.obs"))[1:3, -c(1:3)]
# matplot(t(indexcor), type = "l", col = 9, lty = 1)




# clustering the columns (variables) rather than observations
D2 <- as.matrix(1-abs(cor(tr)))

# image(D2)


h2<- hclust(as.dist(D2), method = "ward.D2")
dnew <- as.dendrogram(h2)
# plot(dnew)
# library(pheatmap)
# pheatmap(tr, cluster_cols = TRUE, clustering_distance_cols = "correlation",  cutree_cols  = 15)
heatmap(D2,
        Colv=as.dendrogram(h2),     
        Rowv=as.dendrogram(h2))


d2 <- color_branches(dnew,k=15, col = 2:16) 
# auto-coloring 4 clusters of branches.
 plot(d2)

# plot(reorder(d2, D2, method = "OLO")) 

labl.cols <- cutree(d2,15) # 15 variables
# matplot(t(tr), type = "l", col = labl.Ave, lty = 1)
# points(1:1060, rep(-5.8, 1060), col = labl.cols)

# image(t(indexcor))
# Pick 15 top features
nonz <- matrix(0,3,15)


for (j in unique(labl.cols)){
  for (i in 1:3){nonz[i, j] <- (which(indexcor[i,]==max(indexcor[i,labl.cols==j])))}
  }
  
  

# matplot(t(tr), type = "l", col = labl.Ave, lty = 1)
# abline(v = nonz[1,], col = 1:15) # pick values in a sensible region

# max(indexcor[1,])

# pairs(cbind(ys_tr[,1], tr[,nonz[1,]]), col = labl.Ave)
```


## Add more ensamble classifiers



* MA: model averaging
* Stack: regression
* Stack: non negative coefficients
* Stack: lasso
* Stack: RF and all the predictors

### Crossvalidation set up

* Split the data into training and testing (200 observations for the training), the rest for testing, 50 random splits. 

* 10 fold cross - validation on each of the 50 training sets to train stack. 

```{r echo = FALSE, message=FALSE, warning= FALSE, eval = FALSE}

N <- 50
grid <- 10^seq(-3, 3, length = 100)
RMSEFULL <- vector(mode = "list", length = 17)
t1 <- vector(mode = "list", length = 3)
flds <- createFolds(1:200, k = 10, list = TRUE, returnTrain = FALSE)


  pls.pc <- c(4,3,4)
  lmfitSG <- replicate(3, vector(mode = "list", length = N))
  lmfitSG2<- replicate(3, vector(mode = "list", length = N))
  rf.fitSG <- replicate(3, vector(mode = "list", length = N)) 
  lmfitSG3 <- replicate(3, vector(mode = "list", length = N)) 
  
  predictions.te3 <- array(0,c(nrow(tr)-200,12, N))
  predictions.tv3 <- array(0,c(200,12, N))

  for(yind in 1:3){
  dat <- cbind(ys_tr[,yind], tr)
  
  dat <- as.data.frame(dat)
  p<- dat %>% ggplot(aes(x= V1)) + geom_histogram()
  print(p)
  dat <- dat %>% na.omit()
  
  nonzy <- nonz[yind,]
  ttt <- replicate(N, sample(nrow(dat), 200))
 
  
  RMSE <- matrix(0, N, 17)
  predictions.te <- matrix(0, nrow(dat)-200, 12)
  predictions.tv <- matrix(0, 200, 12)
 


  for (i in 1:N){


    dat.tr.full <- dat[ttt[,i],]
    dat.te <- dat[-ttt[,i],]
 
    for(l in 1:10){
        dat.tr <- dat.tr.full[-flds[[l]],]
        dat.tv <- dat.tr.full[flds[[l]],]
        
        j <- 1
        x <- model.matrix(V1 ~. , data = dat.tr)
        y1 <- dat.tr[,1]
        lasso.fit <- glmnet(x,y1,alpha=1, lambda = grid) # for lasso
        cv.out <- cv.glmnet(x,y1,alpha=1)
        # print(cv.out$lambda.min)
        
        lasso.fit <- glmnet(x,y1,alpha=1, 
                            lambda = cv.out$lambda.min)
        
        
        lasso.pred <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.tv))
        predictions.tv[flds[[l]] ,j] <- lasso.pred 
        j <- j + 1
        
        #############################################
        
        en.fit <- glmnet(x,y1,alpha=0.5, lambda = grid) # for lasso
        
        # plot(lasso.fit)
        
        cv.out <- cv.glmnet(x,y1,alpha=0.5)
        # print(cv.out$lambda.min)
        
        en.fit <- glmnet(x,y1,alpha=0.5, 
                         lambda = cv.out$lambda.min)
        
        
        en.pred <- predict(en.fit, newx = model.matrix(V1 ~. , data = dat.tv))
        predictions.tv[flds[[l]] ,j] <- en.pred 
        j <- j + 1
        
        
        # cbind(coef(lasso.fit), coef(en.fit))
        ####################################################
        
        pc <- prcomp(dat.tr[, 2:ncol(dat.tr)], scale = TRUE)
        
        lmfit<- lm(dat.tr[, 1]~pc$x[,1:5])
        
        x.tv <- predict(pc, dat.tv[, 2:ncol(dat.tr)])
        # pca.pred.poor <- as.matrix(cbind(rep(1, 200), pc$x[,1:5])) %*% as.matrix(lmfit$coef[1:6])
        pca.pred.poor.te <- as.matrix(cbind(rep(1, nrow(x.tv)), x.tv[,1:5])) %*% as.matrix(lmfit$coef[1:6])
        
        predictions.tv[flds[[l]] ,j] <- pca.pred.poor.te 
        j <- j + 1
        # plot(dat.tv[,1], pca.pred.poor.te, col = as.numeric(dat.tv$class), main = "PCA")
        # abline(a = 0, b = 1)
        
        #############################################################
        
        
        pls.fit <- plsr(V1 ~., data=dat.tr, scale= TRUE, validation = "CV")
        
        
        pls.pred <- predict(pls.fit, dat.tv[,-1], ncomp = pls.pc[yind])
        
        predictions.tv[flds[[l]] ,j] <- pls.pred 
        j <- j+ 1
        
   
        
        #############################################################
        
        lmfit <- lm(dat.tr[, 1]~ as.matrix(dat.tr[, nonzy+1]) )
        x <- model.matrix(V1 ~. , data = dat.tv[, c(1, nonzy+1)])
        pca.pred.poor.te <- x %*% as.matrix(lmfit$coef)
        
        
        predictions.tv[flds[[l]] ,j] <- pca.pred.poor.te 
        j <- j+ 1
        
        #############################################################
        
        
        
        rf.fit <- ranger(V1 ~ ., data = dat.tr, importance = "impurity")
        # plot(rf.fit$variable.importance)
        pred.rf <- predict(rf.fit , data = dat.tv)
        
        predictions.tv[flds[[l]] ,j] <- pred.rf$predictions 
        j <- j+1
        
        
        rf.fit <- ranger(V1 ~ ., data = dat.tr, importance = "impurity",regularization.factor = 0.2, regularization.usedepth=FALSE)
        # plot(rf.fit$variable.importance)
        pred.rf <- predict(rf.fit , data = dat.tv)
        
        predictions.tv[flds[[l]] ,j] <- pred.rf$predictions
        j <- j+1
        # plot(dat.tv[,1], pred.rf$predictions, col = as.numeric(dat.tv$class))
        # abline(0,1)
        
        
        
        bart_machine = bartMachine(as.data.frame(dat.tr[,-1]), dat.tr[,1], verbose = FALSE)
        # summary(bart_machine)
        pred.bart <- predict(bart_machine, as.data.frame(dat.tv[,-1]))
        predictions.tv[flds[[l]] ,j] <- pred.bart
        j <-j+1
        # plot(dat.tv[,1], pred.bart, col = as.numeric(dat.tv$class), main = "BART")
        # abline(0,1)
        
        
        Ktrain <- kernelMatrix(kfunc, as.matrix(dat.tr[,2:ncol(dat.tr)]))
        Kvalid <- kernelMatrix(kfunc, as.matrix(dat.tv[,2:ncol(dat.tr)]), as.matrix(dat.tr[,2:ncol(dat.tr)]))
        kpcTrain <- kPCA(Ktrain)
        
        
        # plot the data projection on the principal components
        
        # pairs(cbind(dat.tr[,1], kpcTrain$KPCs[ , 1 : 6]), col = dat.tr$class)
        kpcValid <- predict(kpcTrain, Kvalid)
        # pairs(kpcTest[ , 1 : 6], col = dat.tv$class)
        lmfit <- lm(dat.tr[, 1]~  kpcTrain$KPCs[ , 1 : 6] )
        
        pca.pred.poor.te <- cbind(rep(1,nrow(kpcValid)), kpcValid[ , 1 : 6]) %*% as.matrix(lmfit$coef)
        # plot(dat.tv[,1], pca.pred.poor.te , col = as.numeric(dat.tv$class), main = "KPCs")
        # abline(0,1)
        
        predictions.tv[flds[[l]] ,j] <- pca.pred.poor.te
        j <- j+ 1
        
        
        
        
        rf.fit <- ranger(y = dat.tr[, 1], x =  kpcTrain$KPCs[ , 1 : 6], importance = "impurity")
        # plot(rf.fit$variable.importance)
        pred.rf <- predict(rf.fit , data = kpcValid[ , 1 : 6])
        predictions.tv[flds[[l]] ,j] <- pred.rf$predictions
        j <- j+1
        
        
        
        rf.fit <- ranger(y = dat.tr[, 1], x =  Ktrain, importance = "impurity")
        # plot(rf.fit$variable.importance)
        pred.rf <- predict(rf.fit , data = Kvalid)
        predictions.tv[flds[[l]] ,j] <- pred.rf$predictions
        j <- j+1
        
        
        
        modelsvm <- svm(V1~.,dat.tr)
        
        #Predict using SVM regression
        predYsvm = predict(modelsvm, dat.tv)
        predictions.tv[flds[[l]] ,j] <- predYsvm
        j <- j+1
        # plot(dat.tv[,1], predYsvm,  main = "RF")
        # abline(0,1)
        
    
    }
    
     
    colnames(predictions.tv) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
    # pairs(cbind(dat.tr.full[,1], predictions.tv)) # correlated predictions
    
    predictions.tv3[,,i] <- predictions.tv
    
    
    lmfitSG[[i]][[yind]]<- lm(dat.tr.full[,1]~predictions.tv - 1) 
    
    lmfitSG2[[i]][[yind]] <-  glmnet(predictions.tv,dat.tr.full[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
    
  
    
  
  
    # b <- glmnet(predictions.tv,dat.tr.full[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE)
    # coef(b)
    # 
    # for(i in 1:50) print(sum(coef(glmnet(predictions.tv3[,,i] ,dat[ttt[,i],1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE))[2:13] - coef(nnls(predictions.tv3[,,i] ,dat[ttt[,i],1]))))
    # 
    # d <- nnls(predictions.tv, dat.tr.full[,1])
    # coef(d)
    # 
    # # summary(  lmfit)
    # 
     # coef(lmfitSG2[[34]][[3]])
   #   i<- 34
   #  dat[ttt[,i],]
   # for(i in 1:50) print(sum(coef(glmnet(predictions.tv3[,,i] ,dat[ttt[,i],1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE))))
   #  
   #  failed <- rep(0,50)
   #  for(i in 1:50)   failed[i] <- round(sum(coef( lmfitSG2[[i]][[3]])),2)
   #  b<- cv.glmnet(predictions.tv3[,,i] ,dat[ttt[,i],1],alpha=1, lower.limits = 0, intercept = FALSE)
   #  cbind(failed, RMSE[,17])

    rf.fitSG[[i]][[yind]] <- ranger(y = dat.tr.full[, 1], x = cbind( predictions.tv, dat.tr.full[, -1]), importance = "impurity")
    
        cv.out <- cv.glmnet(predictions.tv,dat.tr.full[,1],alpha=1, lambda = grid,  intercept = FALSE)
        
       lmfitSG3[[i]][[yind]] <- glmnet(predictions.tv,dat.tr.full[,1],alpha=1, lambda = cv.out$lambda.min,   intercept = FALSE) 
    
    ##-----------------------------------------------------------
    j <- 1
    x <- model.matrix(V1 ~. , data = dat.tr.full)
    y1 <- dat.tr.full[,1]
    lasso.fit <- glmnet(x,y1,alpha=1, lambda = grid) # for lasso
    cv.out <- cv.glmnet(x,y1,alpha=1)
    # print(cv.out$lambda.min)
    
    lasso.fit <- glmnet(x,y1,alpha=1, 
                        lambda = cv.out$lambda.min)
    
    
    lasso.pred <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.te))
    RMSE[i,j] <- calcRMSE(dat.te[,1], lasso.pred)
    predictions.te[,j] <- lasso.pred 
    j <- j + 1
    
    #############################################
    
    en.fit <- glmnet(x,y1,alpha=0.5, lambda = grid) # for lasso
    
    # plot(lasso.fit)
    
    cv.out <- cv.glmnet(x,y1,alpha=0.5)
    # print(cv.out$lambda.min)
    
    en.fit <- glmnet(x,y1,alpha=0.5, 
                     lambda = cv.out$lambda.min)
    
    
    en.pred <- predict(en.fit, newx = model.matrix(V1 ~. , data = dat.te))
    RMSE[i,j] <- calcRMSE(dat.te[,1], en.pred)
    predictions.te[,j] <- en.pred 
    j <- j + 1
    
    ####################################################
    
    pc <- prcomp(dat.tr.full[, 2:ncol(dat.tr.full)], scale = TRUE)
    
    lmfit<- lm(dat.tr.full[, 1]~pc$x[,1:5])
    
    x.te <- predict(pc, dat.te[, 2:ncol(dat.tr.full)])
    # pca.pred.poor <- as.matrix(cbind(rep(1, 200), pc$x[,1:5])) %*% as.matrix(lmfit$coef[1:6])
    pca.pred.poor.te <- as.matrix(cbind(rep(1, nrow(x.te)), x.te[,1:5])) %*% as.matrix(lmfit$coef[1:6])
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
    predictions.te[,j] <- pca.pred.poor.te 
    j <- j + 1
    # plot(dat.te[,1], pca.pred.poor.te, col = as.numeric(dat.te$class), main = "PCA")
    # abline(a = 0, b = 1)
    
    #############################################################
    
    
    pls.fit <- plsr(V1 ~., data=dat.tr.full, scale= TRUE, validation = "CV")
    
    
    pls.pred <- predict(pls.fit, dat.te[,-1], ncomp = pls.pc[yind])
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
    predictions.te[,j] <- pls.pred 
    j <- j+ 1
    

    
    #############################################################
    
    lmfit <- lm(dat.tr.full[, 1]~ as.matrix(dat.tr.full[, nonzy+1]) )
    x <- model.matrix(V1 ~. , data = dat.te[, c(1, nonzy+1)])
    pca.pred.poor.te <- x %*% as.matrix(lmfit$coef)
    
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
    predictions.te[,j] <- pca.pred.poor.te 
    j <- j+ 1
    
    #############################################################
    
    
    
    rf.fit <- ranger(V1 ~ ., data = dat.tr.full, importance = "impurity")
    # plot(rf.fit$variable.importance)
    pred.rf <- predict(rf.fit , data = dat.te)
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
    predictions.te[,j] <- pred.rf$predictions 
    j <- j+1
    
    
    rf.fit <- ranger(V1 ~ ., data = dat.tr.full, importance = "impurity",regularization.factor = 0.2, regularization.usedepth=FALSE)
    # plot(rf.fit$variable.importance)
    pred.rf <- predict(rf.fit , data = dat.te)
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
    predictions.te[,j] <- pred.rf$predictions
    j <- j+1
    # plot(dat.te[,1], pred.rf$predictions, col = as.numeric(dat.te$class))
    # abline(0,1)
    
    
    
    bart_machine = bartMachine(as.data.frame(dat.tr.full[,-1]), dat.tr.full[,1], verbose = FALSE)
    # summary(bart_machine)
    pred.bart <- predict(bart_machine, as.data.frame(dat.te[,-1]))
    RMSE[i,j] <- calcRMSE(dat.te[,1], pred.bart)
    predictions.te[,j] <- pred.bart
    j <-j+1
    # plot(dat.te[,1], pred.bart, col = as.numeric(dat.te$class), main = "BART")
    # abline(0,1)
    
    
    Ktrain <- kernelMatrix(kfunc, as.matrix(dat.tr.full[,2:ncol(dat.tr.full)]))
    Ktest <- kernelMatrix(kfunc, as.matrix(dat.te[,2:ncol(dat.tr.full)]), as.matrix(dat.tr.full[,2:ncol(dat.tr.full)]))
    kpcTrain <- kPCA(Ktrain)
    
    
    # plot the data projection on the principal components
    
    # pairs(cbind(dat.tr.full[,1], kpcTrain$KPCs[ , 1 : 6]), col = dat.tr.full$class)
    kpcTest <- predict(kpcTrain, Ktest)
    # pairs(kpcTest[ , 1 : 6], col = dat.te$class)
    lmfit <- lm(dat.tr.full[, 1]~  kpcTrain$KPCs[ , 1 : 6] )
    
    pca.pred.poor.te <- cbind(rep(1,nrow(kpcTest)), kpcTest[ , 1 : 6]) %*% as.matrix(lmfit$coef)
    # plot(dat.te[,1], pca.pred.poor.te , col = as.numeric(dat.te$class), main = "KPCs")
    # abline(0,1)
    
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
    predictions.te[,j] <- pca.pred.poor.te
    j <- j+ 1
    
    
    
    
    rf.fit <- ranger(y = dat.tr.full[, 1], x =  kpcTrain$KPCs[ , 1 : 6], importance = "impurity")
    # plot(rf.fit$variable.importance)
    pred.rf <- predict(rf.fit , data = kpcTest[ , 1 : 6])
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
    predictions.te[,j] <- pred.rf$predictions
    j <- j+1
    
    
    
    rf.fit <- ranger(y = dat.tr.full[, 1], x =  Ktrain, importance = "impurity")
    # plot(rf.fit$variable.importance)
    pred.rf <- predict(rf.fit , data = Ktest)
    
    RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
    predictions.te[,j] <- pred.rf$predictions
    j <- j+1
    
    
    
    modelsvm <- svm(V1~.,dat.tr.full)
    
    #Predict using SVM regression
    predYsvm = predict(modelsvm, dat.te)
    RMSE[i,j] <- calcRMSE(dat.te[,1], predYsvm)
    predictions.te[,j] <- predYsvm
    j <- j+1
    # plot(dat.te[,1], predYsvm,  main = "RF")
    # abline(0,1)
    
    
    
    RMSE[i,j] <-  calcRMSE(dat.te[,1],   rowMeans(predictions.te[, 1:12]))
 

    colnames(predictions.te) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
    # pairs(predictions.te)
    
    predictions.te3[1:nrow(predictions.te),,i] <- predictions.te
     
  pred.stack <- as.matrix( predictions.te) %*% as.matrix(lmfitSG[[i]][[yind]]$coef)
 
 pred.stack2 <- predict(lmfitSG2[[i]][[yind]], predictions.te)
     # predict(b, lambda =0, predictions.te3[1:nrow(predictions.te),,i] )
# coef(b, lambda =0)
  pred.stackrf <- predict(rf.fitSG[[i]][[yind]], data = cbind(predictions.te, dat.te[, -1]))$prediction

  
   pred.stack3 <- predict(lmfitSG3[[i]][[yind]], predictions.te)
   
  # plot(pred.stack2, pred.stack)
# summary(  lmfit)
#   i <- 38
#  plot(dat.te[,1], predict(lmfitSG2[[i]][[yind]], predictions.te3[1:nrow(predictions.te),,i] ))
# sum(coef(lmfitSG2[[38]][[yind]]))
# 
# calcRMSE(dat.te[,1], predict(lmfitSG2[[i]][[yind]],predictions.te3[1:nrow(predictions.te),,i]  ))
# 
# coef(lmfitSG2[[i]][[yind]])
# RMSEtemp <- seq(0,50)
# for(i in 1:50)RMSEtemp[i] <- calcRMSE(dat[-ttt[,i],1], predict(lmfitSG2[[i]][[yind]],predictions.te3[1:nrow(predictions.te),,i]  ))
# cbind(RMSEtemp[1:50], RMSE[, 17] )
# RMSE[, 17] <- RMSEtemp[1:50]
# 
# RMSE[i, 17] <- calcRMSE(dat.te[,1], predict(lmfitSG2[[i]][[yind]],predictions.te3[1:nrow(predictions.te),,i]  ))
#  failed
# pairs(predictions.te3[1:nrow(predictions.te),,i] )
   RMSE[i,j+1] <- calcRMSE(dat.te[,1],  pred.stack)
   RMSE[i,j+2] <- calcRMSE(dat.te[,1],  pred.stack2)
   RMSE[i,j+3] <- calcRMSE(dat.te[,1],  pred.stackrf)
   RMSE[i,j+4] <- calcRMSE(dat.te[,1],  pred.stack3)
  }
    
  colnames(RMSE) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm", "ensamble MA",  "ensLM",  "ensNonNeg", "ensRF", "ensLasso")
  RMSEFULL[[yind]] <- RMSE
  
  RMSE1 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:17,"ALGORITHM")
  # RMSE1 %>% ggplot(aes(x = ALGORITHM, y = value)) + geom_boxplot()
  p1<- RMSE1 %>% filter(ALGORITHM!="ensNonNeg") %>% group_by(ALGORITHM) %>% mutate(meanv = mean(value)) %>% ggplot(aes(x= ALGORITHM, y = value, color = split, group = split))+
    geom_point()+ geom_line() +geom_line(aes(x = ALGORITHM, y = meanv), col = 1)+ theme(legend.position = "none")+
    ylab("RMSE")+ ggtitle(names(ys_tr)[yind])+
    theme(axis.text.x=element_text(angle = 90, hjust = 0.5, vjust = 0.5))
  
  print(p1)
  # t1
  #     RMSE[failed==0, 17] <- NA
  #     RMSE[failed==1, 17] 
  #     
  #     RMSEFULL[[yind]]  <- RMSE
  #      RMSE1 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:18,"ALGORITHM")
  # 
       t1[[yind]] <- RMSE1 %>% group_by(ALGORITHM) %>% summarise(meanv = mean(value, na.rm = TRUE), sdv=sd(value, na.rm = TRUE)) %>% arrange(desc(meanv))
  
}







save.image("analysisEnsambles3b.Rdata")

```

```{r echo = FALSE}

load("analysisEnsambles3b.Rdata")

RMSEFULLDF <- matrix(0, 850*3, 3) %>%as.data.frame()%>% mutate(trait = rep(names(ys),each = 850))
# glimpse(RMSEFULLDF)

# plot(RMSEFULLDF$Split[1:900])

for(yind in 1:3){  
  a <- (yind-1)*850+1
  b <- yind*850
  d<- (yind-1)*50+1
  e<- yind*50
  RMSE <- RMSEFULL[[yind]]  
  RMSE1 <- RMSE  %>%as.data.frame()%>% mutate(split = d:e)%>% pivot_longer(1:17,"ALGORITHM")
  RMSEFULLDF[a:b,1:3]<- RMSE1
}
names(RMSEFULLDF) <- c("Split", "Algorithm", "RMSE", "Trait")

RMSEFULLDF <- RMSEFULLDF%>% mutate(Split = as.factor(Split), Algorithm = as.factor(Algorithm), Trait = as.factor(Trait))

# data.nest1.aov <- aov(RMSE ~ Algorithm *Trait + Split, RMSEFULLDF)
# anova(data.nest1.aov)
# TukeyHSD(data.nest1.aov)

```

```{r echo = TRUE}


data.nest1.lmer <- lmer(RMSE ~ Algorithm*Trait + (1 | Split), RMSEFULLDF, REML = TRUE)

print(summary(data.nest1.lmer)[13])

```

```{r echo = FALSE}
# plot(allEffects(data.nest1.lmer), axes = list(x=list(rotate = 90)))

data.nest1.eff <- as.data.frame(effect("Algorithm:Trait", data.nest1.lmer))
data.nest1.eff <- data.nest1.eff %>% mutate(nonneg = as.numeric(Algorithm=="ensNonNeg"))

# data.nest1.eff$nonneg
# glimpse(data.nest1.eff )
data.nest1.eff %>% ggplot( aes(y = fit, x = Trait, group = Algorithm, colour = Algorithm)) + geom_pointrange(aes(ymin = lower,
    ymax = upper)) + scale_y_continuous("RMSE") + theme(axis.text.x = element_text(angle = 0))+geom_line(aes(alpha = nonneg), show.legend = FALSE)
```

```{r echo = FALSE, fig.height=14, fig.width = 10}
data.nest1.eff %>% ggplot( aes(y = fit, x =  Algorithm)) + geom_errorbar(aes(ymin = lower,
    ymax = upper, colour = nonneg), show.legend = FALSE) + scale_y_continuous("RMSE") + theme(axis.text.x = element_text(angle = 90))+
  facet_grid(Trait ~ .) 

data.nest1.eff <- as.data.frame(Effect("Algorithm", data.nest1.lmer))
```

```{r echo = FALSE}
data.nest1.eff %>% ggplot( aes(y = fit, x = Algorithm)) + geom_pointrange(aes(ymin = lower,
    ymax = upper)) + scale_y_continuous("RMSE") + theme(axis.text.x = element_text(angle = 90)) + geom_point(size = 3, pch = 1)

for(yind in 1:3){
  
    dat <- cbind(ys_tr[,yind], tr)
  
  dat <- as.data.frame(dat)
  p<- dat %>% ggplot(aes(x= V1)) + geom_histogram() + ggtitle(names(ys_tr)[yind])
  print(p)
  
  RMSE <- RMSEFULL[[yind]]  
  RMSE1 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:17,"ALGORITHM")
  # RMSE1 %>% ggplot(aes(x = ALGORITHM, y = value)) + geom_boxplot()
  p1<- RMSE1 %>% group_by(ALGORITHM) %>% mutate(meanv = mean(value, na.rm = TRUE)) %>% ggplot(aes(x= ALGORITHM, y = value, color = split, group = split))+
    geom_point()+ geom_line() +geom_line(aes(x = ALGORITHM, y = meanv), col = 1)+ theme(legend.position = "none")+
    ylab("RMSE")+ ggtitle(names(ys_tr)[yind])+
    theme(axis.text.x=element_text(angle = 90, hjust = 0.5, vjust = 0.5))
  
  print(p1)
  
  LMcoef <- matrix(0,50, 12)
  for (i in 1:50)LMcoef[i,] <- lmfitSG[[i]][[yind]]$coefficients
  
  colnames(LMcoef) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
  
  matplot(t(LMcoef), type = "l", lty = 1, xaxt = "n", col = 9, ylab = "coef", main = "ensemble LM")
  abline(h  = 0)
  axis(1, at = 1 : 12, labels = colnames(LMcoef),las = 2, cex.axis=0.7)
  
   # hist(rowSums(LMcoef))
  
  LMcoef <- matrix(0,50, 12)
  for (i in 1:50)LMcoef[i,] <- coef(lmfitSG2[[i]][[yind]])[2:13]
  colnames(LMcoef) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
  
  matplot(t(LMcoef), type = "l", lty = 1, xaxt = "n", col = 9, ylab = "coef", main = "ensemble Non Neg")
  abline(h  = 0)
  axis(1, at = 1 : 12, labels = colnames(LMcoef),las = 2, cex.axis=0.7)
  
   # hist(rowSums(LMcoef))
  
   LMcoef <- matrix(0,50, 12)
  for (i in 1:50)LMcoef[i,] <- coef(lmfitSG3[[i]][[yind]])[2:13]
  colnames(LMcoef) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
  
  matplot(t(LMcoef), type = "l", lty = 1, xaxt = "n", col = 9, ylab = "coef", main = "ensemble Lasso")
  abline(h  = 0)
  axis(1, at = 1 : 12, labels = colnames(LMcoef),las = 2, cex.axis=0.7)
  
  
# hist(rowSums(LMcoef))
  
  LMcoef <- matrix(0,50, 12)
  for (i in 1:50)LMcoef[i,] <- rf.fitSG[[i]][[yind]]$variable.importance[1:12]
  colnames(LMcoef) <- c("lasso", "elastic net", "pca", "pls", "LM+15", "RF", "RF+VS", "Bart", "kPCA","RF+kPCA", "RF+kernel", "svm")
  
  matplot(t(LMcoef), type = "l", lty = 1, xaxt = "n", col = 9, ylab = "Variable importance", main = "ensemble RF")
  abline(h  = 0)
  axis(1, at = 1 : 12, labels = colnames(LMcoef),las = 2, cex.axis=0.7)
  

  
  # print(t1[yind])
  

  RMSE1 <- RMSE1 %>% mutate(Split = as.factor(split), Algorithm = as.factor(ALGORITHM))
# glimpse(RMSE1)
# b <- aov(value~ split+ ALGORITHM, data = RMSE1)
# anova(b) %>% print()



data.nest1.lmer <- lmer(value ~ Algorithm + (1 | Split), RMSE1, REML = TRUE)

print(summary(data.nest1.lmer)[13])


data.nest1.eff <- as.data.frame(Effect("Algorithm", data.nest1.lmer))



p <- data.nest1.eff %>% ggplot( aes(y = fit, x = Algorithm)) + geom_pointrange(aes(ymin = lower,
    ymax = upper)) + scale_y_continuous("RMSE") + theme(axis.text.x = element_text(angle = 90)) + geom_point(size = 3, pch = 1)+ ggtitle(names(ys_tr)[yind])

print(p)

# glimpse(RMSE1)
# b <- aov(value~ split+ ALGORITHM, data = RMSE1)
# anova(b) %>% print()
# TukeyHSD(b)$ALGORITHM[,4][TukeyHSD(b)$ALGORITHM[,4]<0.05] %>% t() %>%t()

# plot(TukeyHSD(b))

}

```

