---
title: "Vista Milk: predictions from an ensemble model"
author: Katarina Domijan
date: April 29, 2021
output: beamer_presentation
editor_options: 
  chunk_output_type: console
---


## Data:

```{r echo = FALSE, message=FALSE, warning=FALSE, results = 'hide'}
rm(list = ls())
set.seed(1979)
library(tidyverse)
library(dendextend)
library(glmnet) # for the ridge regression
library(pls)
library(ranger)
library(bartMachine)
library(e1071)
library(BKPC)
library(kernlab)
tr <- read.csv("tr.csv") 
te <- read.csv("te.csv") 

calcRMSE <- function(y, yhat){sqrt(mean((yhat - y)^2))}


y <- tr[,1:3]
tr <- tr[, -c(1:3)]


a <- apply(tr,2,mean)
b <- apply(tr,2,sd)


tr <- scale(tr,a,b)
te <- scale(te,a,b)


kfunc <-  rbfdot(sigma = 0.0005)

```



- The data were scaled and centered

```{r fig.cap="Scaled training data", echo=FALSE}
###########################################
# Look for clusters in the training data


D <- as.matrix(dist(tr,method="euclidean"))


h <- hclust(as.dist(D), method = "ward.D2")
d <- as.dendrogram(h)


# plot(d)

# plot(d)
# library(circlize)
# 
# # create a dendrogram
# 
# 
# # modify the dendrogram to have some colors in the branches and labels
d2 <- color_branches(d,k=7, col=c(2,3,5,4, 6,7,8)) 

par(mfrow = c(1, 2))
plot(d2)
# 
# # plot the radial plot
# par(mar = rep(0,4))
# plot(d, labels=FALSE)


labl.Ave <- cutree(d,7)
matplot(t(tr), type = "l", col = labl.Ave+1, lty = 1, ylab = "transmittance values") # see the 7 clusters


dev.off();
```



## Cross - validation set up:

- Split the full training set into 50 random splits of sample size 200 (training) and 353 (validation).
- Fit the models to the randomly selected training (sub)-sets and compare using RMSE of the validation sets.



## Algorithms

Submitted predictions are from an ensemble model -  averages taken over predictions from the following models:


1. pls with 3, 4, and 5 components (R package **pls**)
2. linear regression with 6 PCs with highest eigenvalues
3. lasso (R package **glmnet**)
4. random forest (R package **ranger**) with and without regularization
5. BART (R package **bartMachine**)

## Classifiers cont'd

6. support vector machine for regression (R package **e1071**)
7. linear regression with 6 KPCs with highest eigenvalues using Gaussian kernel (R package **kernlab** and **BKPC**)
8. random forest with 6 KPCs with highest eigenvalues using Gaussian kernel
9. random forest with a Gaussian kernel
10. linear regression with 15 wavelengths selected by a) clustering the wavelengths into 15 clusters and selecting the wavelength from each cluster that has the highest correlation with the response variable


```{r echo=FALSE}
# clustering the columns (variables) rather than observations
D2 <- as.matrix(1-abs(cor(tr)))

# image(D)


h2<- hclust(as.dist(D2), method = "ward.D2")
dnew <- as.dendrogram(h2)


indexcor <- abs(cor(cbind(y,tr), use = "pairwise.complete.obs"))[1:3, -c(1:3)]



labl.cols <- cutree(dnew,15) # 15 variables


# Pick 15 top features
nonz1 <- NULL
nonz2 <- NULL
nonz3 <- NULL
for (j in unique(labl.cols)){
  nonz1[j] <- (which(indexcor[1,]==max(indexcor[1,labl.cols==j])))
  nonz2[j] <- (which(indexcor[2,]==max(indexcor[2,labl.cols==j])))
  nonz3[j] <- (which(indexcor[3,]==max(indexcor[3,labl.cols==j])))


  }
```

## kappa_casein

```{r echo = FALSE}
N <- 50
```


```{r, fig.cap="kappa_casein", echo=FALSE, message=FALSE, warning=FALSE}


dat <- cbind((y[,1]), tr)
dat <- as.data.frame(dat)
dat <- dat %>% mutate(class = as.factor(labl.Ave))
dat <- dat %>% na.omit()
nonz <- nonz1

ttt <- replicate(N, sample(nrow(dat), 130+130))
ttr <- ttt[1:130,]
ttv <- ttt[131:260,]

RMSE <- matrix(0, N, 17)
predictions.te <- matrix(0,nrow(dat)-260, 13)
predictions.tv <- matrix(0,130, 13)

Coefs1 <- matrix(0,N, 13)
Coefs2 <- matrix(0,N, 13)
Coefs3 <- matrix(0,N, 13)
i <- 1
for (i in 1:N){
  j <- 1
  dat.tr <- dat[ttr[,i],]
  dat.tv <- dat[ttv[,i],]
  dat.te <- dat[-ttt[,i],]


  x <- model.matrix(V1 ~. , data = dat.tr[, -1062])

  y1 <- dat.tr[,1]

  grid <- 10^seq(-3, 3, length = 100)




  lasso.fit <- glmnet(x,y1,alpha=1, lambda = grid) # for
  cv.out <- cv.glmnet(x,y1,alpha=1)


  lasso.fit <- glmnet(x,y1,alpha=1,
                      lambda = cv.out$lambda.min)



  lasso.pred <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.te[, -1062]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], lasso.pred)
  predictions.te[,j] <- lasso.pred
  predictions.tv[,j] <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.tv[, -1062]))
  j <- j + 1


###########################################




  pc <- prcomp(dat.tr[, 2:1061], scale = TRUE)



  lmfit<- lm(dat.tr[, 1]~pc$x[,1:5])

  x.te <- predict(pc, dat.te[, 2:1061])

  pca.pred.poor.te <- as.matrix(cbind(rep(1, nrow(x.te)), x.te[,1:5])) %*% as.matrix(lmfit$coef[1:6])
  
  x.tv <- predict(pc, dat.tv[, 2:1061])

  pca.pred.poor.tv <- as.matrix(cbind(rep(1, nrow(x.tv)), x.tv[,1:5])) %*% as.matrix(lmfit$coef[1:6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j + 1


###########################################
  pls.fit <- plsr(V1 ~., data=dat.tr[, -1062], scale= TRUE, validation = "CV")

  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 3)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
  predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 3)
  j <- j+ 1

  
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 4)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 4)
  j <- j+ 1
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 5)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 5)
  j <- j+ 1

  
  ###########################################
  lmfit <- lm(dat.tr[, 1]~ as.matrix(dat.tr[, nonz+1]) )


  x <- model.matrix(V1 ~. , data = dat.te[, c(1, nonz+1)])
  pca.pred.poor.te <- x %*% as.matrix(lmfit$coef)
 pca.pred.poor.tv <- model.matrix(V1 ~. , data = dat.tv[, c(1, nonz+1)]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

  ###########################################


  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity")

  pred.rf <- predict(rf.fit , data = dat.te[, -1062])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions 
  j <- j+1


###########################################
  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity",regularization.factor = 0.2, regularization.usedepth=FALSE)
  # plot(rf.fit$variable.importance)
  pred.rf <- predict(rf.fit , data = dat.te)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
    predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions
  j <- j+1

###########################################


  bart_machine = bartMachine(as.data.frame(dat.tr[,-c(1,1062)]), dat.tr[,1], verbose = FALSE)
  # summary(bart_machine)
  pred.bart <- predict(bart_machine, as.data.frame(dat.te[,-c(1,1062)]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.bart)
  predictions.te[,j] <- pred.bart
    predictions.tv[,j] <- predict(bart_machine, as.data.frame(dat.tv[,-c(1, 1062)]))
  j <-j+1

###########################################

  Ktrain <- kernelMatrix(kfunc, as.matrix(dat.tr[,2:1061]))
  Ktest <- kernelMatrix(kfunc, as.matrix(dat.te[,2:1061]), as.matrix(dat.tr[,2:1061]))
   Kvalid <- kernelMatrix(kfunc, as.matrix(dat.tv[,2:1061]), as.matrix(dat.tr[,2:1061]))
 
  kpcTrain <- kPCA(Ktrain)


  kpcTest <- predict(kpcTrain, Ktest)
  kpcValid <- predict(kpcTrain, Kvalid)
  # pairs(kpcTest[ , 1 : 6], col = dat.te$class)
  lmfit <- lm(dat.tr[, 1]~  kpcTrain$KPCs[ , 1 : 6] )





  pca.pred.poor.te <- cbind(rep(1,nrow(kpcTest)), kpcTest[ , 1 : 6]) %*% as.matrix(lmfit$coef)

  pca.pred.poor.tv <- cbind(rep(1,nrow(kpcValid)), kpcValid[ , 1 : 6]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

###########################################


  rf.fit <- ranger(y = dat.tr[, 1], x =  kpcTrain$KPCs[ , 1 : 6], importance = "impurity")

  pred.rf <- predict(rf.fit , data = kpcTest[ , 1 : 6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = kpcValid[ , 1 : 6])$predictions
  j <- j+1

###########################################

  rf.fit <- ranger(y = dat.tr[, 1], x =  Ktrain, importance = "impurity")

  pred.rf <- predict(rf.fit , data = Ktest)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = Kvalid)$predictions
  j <- j+1

###########################################

  modelsvm <- svm(V1~.,dat.tr[, -1062])

  #Predict using SVM regression
  predYsvm = predict(modelsvm, dat.te[, -1062])
  RMSE[i,j] <- calcRMSE(dat.te[,1], predYsvm)
  predictions.te[,j] <- predYsvm
  predictions.tv[,j] <- predict(modelsvm, dat.tv[, -1062])
  j <- j+1

########################################### 

  RMSE[i,j] <-  calcRMSE(dat.te[,1],   rowMeans(predictions.te[, 1:13]))

    
   colnames(predictions.tv) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    colnames(predictions.te) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    
  lmfit<- lm(dat.tv[, 1]~predictions.tv[, 1:13] - 1)
# summary(  lmfit)
  
  pred.stack <- as.matrix( predictions.te[, 1:13]) %*% as.matrix(lmfit$coef)
  # pred.stack <- as.matrix(cbind(rep(1, nrow(x.te)), predictions.te[, 1:13])) %*% as.matrix(lmfit$coef)
   lmfit2<- glmnet(predictions.tv,dat.tv[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
# summary(  lmfit)
  
    pred.stack2 <- predict(lmfit2, predictions.te)
    
    rf.fitSG <- ranger(y = dat.tv[, 1], x = cbind( predictions.tv, dat.tv[, -c(1,1062)]), importance = "impurity")

  pred.rf <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))

# summary(  lmfit)
  
    pred.stack3 <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))$prediction
   RMSE[i,j+1] <- calcRMSE(dat.te[,1],  pred.stack)
    RMSE[i,j+2] <- calcRMSE(dat.te[,1],  pred.stack2)
      RMSE[i,j+3] <- calcRMSE(dat.te[,1],  pred.stack3)
    
  Coefs1[i,] <- coef(lmfit)
  Coefs2[i,] <- coef(lmfit2)[2:14]
  # Coefs3[i,] <- plot(rf.fitSG$variable.importance)
}

 # matplot(t(Coefs3), type = "l")
# matplot((Coefs1), type = "l")
# 
# matplot((Coefs2), type = "l")
# 
# pairs(Coefs2)
# 
# for (k in 1:13)plot(Coefs1[,k], Coefs2[,k])
# 
# image(t(Coefs1))
# image(t(Coefs2))
# plot(dat.te[,1],rowMeans(predictions.te[, 1:13]),  col = as.numeric(dat.te$class)+1, xlab = "kappa_casein validation set: one random split", ylab = "predicted",  main = "ensamble")
# abline(0,1)

# plot(dat.te[,1],pred.stack2,  col = as.numeric(dat.te$class)+1, xlab = "kappa_casein validation set: one random split", ylab = "predicted",  main = "ensamble")
# abline(0,1)

# colnames(predictions.tv) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")

# plot(lmfit$coef, lmfit2$coef)

# dim(dat.te)
# pairs(rbind(cbind(dat.tv[, 1],predictions.tv[, 1:13]), cbind(dat.te[, 1],predictions.te[, 1:13])), col = c(rep(1, 130), rep(2, 139)))
# 
# dattemp<- as.data.frame(cbind(dat.tv[, 1], predictions.tv[, 1:13]))
#    lmfit<- lm(V1~. - 1, data = dattemp)
#    anova(lmfit)




# 
#   lasso.fit <- glmnet(predictions.tv,dat.tv[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
# 
#     lasso.fit <- glmnet(predictions.te,dat.te[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
# 
# coef( lasso.fit)
# 
# library(condvis2)
# condvis(dattemp,  lasso.fit,  response = "V1")
#   
# library(car)
# avPlots(lmfit2, col = 3)

```


<!-- stack 1 : gives more weight to algorithms that are "overfitting" - have better predictions in the training set -->

<!-- stack 2 : gives more weight to algorithms that are better predict the test set. But how will they generalise to the validation set? -->

## kappa_casein



```{r, fig.cap="kappa_casein results", echo=FALSE, message=FALSE, warning=FALSE}
colnames(RMSE) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm",  "ensMA",  "ensLM",  "ensNonNeg", "ensRF")

RMSE1 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:17,"ALGORITHM")

# RMSE1 %>% ggplot(aes(x = ALGORITHM, y = value)) + geom_boxplot()
RMSE1 %>% group_by(ALGORITHM) %>% mutate(meanv = mean(value)) %>% ggplot(aes(x= ALGORITHM, y = value, color = split, group = split))+
  geom_point()+ geom_line() +geom_line(aes(x = ALGORITHM, y = meanv), col = 1)+ theme(legend.position = "none")+
  ylab("RMSE")+
    theme(axis.text.x=element_text(angle = 90, hjust = 0.5, vjust = 0.5))

t1 <- RMSE1 %>% group_by(ALGORITHM) %>% summarise(meanv = mean(value), sdv=sd(value)) %>% arrange(desc(meanv))



```


## kappa_casein

```{r echo = FALSE}
knitr::kable((t1))
```


## Casein_micelle_size



```{r fig.cap="Casein_micelle_size", echo=FALSE, message=FALSE, warning=FALSE}


dat <- cbind(log(y[,2]), tr)
dat <- as.data.frame(dat)
dat <- dat %>% mutate(class = as.factor(labl.Ave))
dat <- dat %>% na.omit()
nonz <- nonz2



# ====replace below]
RMSE <- matrix(0, N, 17)
predictions.te <- matrix(0,nrow(dat)-260, 13)
predictions.tv <- matrix(0,130, 13)

Coefs1 <- matrix(0,N, 13)
Coefs2 <- matrix(0,N, 13)
Coefs3 <- matrix(0,N, 13)
i <- 1
for (i in 1:N){
  j <- 1
  dat.tr <- dat[ttr[,i],]
  dat.tv <- dat[ttv[,i],]
  dat.te <- dat[-ttt[,i],]


  x <- model.matrix(V1 ~. , data = dat.tr[, -1062])

  y1 <- dat.tr[,1]

  grid <- 10^seq(-3, 3, length = 100)




  lasso.fit <- glmnet(x,y1,alpha=1, lambda = grid) # for
  cv.out <- cv.glmnet(x,y1,alpha=1)


  lasso.fit <- glmnet(x,y1,alpha=1,
                      lambda = cv.out$lambda.min)



  lasso.pred <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.te[, -1062]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], lasso.pred)
  predictions.te[,j] <- lasso.pred
  predictions.tv[,j] <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.tv[, -1062]))
  j <- j + 1


###########################################




  pc <- prcomp(dat.tr[, 2:1061], scale = TRUE)



  lmfit<- lm(dat.tr[, 1]~pc$x[,1:5])

  x.te <- predict(pc, dat.te[, 2:1061])

  pca.pred.poor.te <- as.matrix(cbind(rep(1, nrow(x.te)), x.te[,1:5])) %*% as.matrix(lmfit$coef[1:6])
  
  x.tv <- predict(pc, dat.tv[, 2:1061])

  pca.pred.poor.tv <- as.matrix(cbind(rep(1, nrow(x.tv)), x.tv[,1:5])) %*% as.matrix(lmfit$coef[1:6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j + 1


###########################################
  pls.fit <- plsr(V1 ~., data=dat.tr[, -1062], scale= TRUE, validation = "CV")

  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 3)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
  predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 3)
  j <- j+ 1

  
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 4)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 4)
  j <- j+ 1
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 5)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 5)
  j <- j+ 1

  
  ###########################################
  lmfit <- lm(dat.tr[, 1]~ as.matrix(dat.tr[, nonz+1]) )


  x <- model.matrix(V1 ~. , data = dat.te[, c(1, nonz+1)])
  pca.pred.poor.te <- x %*% as.matrix(lmfit$coef)
 pca.pred.poor.tv <- model.matrix(V1 ~. , data = dat.tv[, c(1, nonz+1)]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

  ###########################################


  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity")

  pred.rf <- predict(rf.fit , data = dat.te[, -1062])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions 
  j <- j+1


###########################################
  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity",regularization.factor = 0.2, regularization.usedepth=FALSE)
  # plot(rf.fit$variable.importance)
  pred.rf <- predict(rf.fit , data = dat.te)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
    predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions
  j <- j+1

###########################################


  bart_machine = bartMachine(as.data.frame(dat.tr[,-c(1,1062)]), dat.tr[,1], verbose = FALSE)
  # summary(bart_machine)
  pred.bart <- predict(bart_machine, as.data.frame(dat.te[,-c(1,1062)]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.bart)
  predictions.te[,j] <- pred.bart
    predictions.tv[,j] <- predict(bart_machine, as.data.frame(dat.tv[,-c(1, 1062)]))
  j <-j+1

###########################################

  Ktrain <- kernelMatrix(kfunc, as.matrix(dat.tr[,2:1061]))
  Ktest <- kernelMatrix(kfunc, as.matrix(dat.te[,2:1061]), as.matrix(dat.tr[,2:1061]))
   Kvalid <- kernelMatrix(kfunc, as.matrix(dat.tv[,2:1061]), as.matrix(dat.tr[,2:1061]))
 
  kpcTrain <- kPCA(Ktrain)


  kpcTest <- predict(kpcTrain, Ktest)
  kpcValid <- predict(kpcTrain, Kvalid)
  # pairs(kpcTest[ , 1 : 6], col = dat.te$class)
  lmfit <- lm(dat.tr[, 1]~  kpcTrain$KPCs[ , 1 : 6] )





  pca.pred.poor.te <- cbind(rep(1,nrow(kpcTest)), kpcTest[ , 1 : 6]) %*% as.matrix(lmfit$coef)

  pca.pred.poor.tv <- cbind(rep(1,nrow(kpcValid)), kpcValid[ , 1 : 6]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

###########################################


  rf.fit <- ranger(y = dat.tr[, 1], x =  kpcTrain$KPCs[ , 1 : 6], importance = "impurity")

  pred.rf <- predict(rf.fit , data = kpcTest[ , 1 : 6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = kpcValid[ , 1 : 6])$predictions
  j <- j+1

###########################################

  rf.fit <- ranger(y = dat.tr[, 1], x =  Ktrain, importance = "impurity")

  pred.rf <- predict(rf.fit , data = Ktest)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = Kvalid)$predictions
  j <- j+1

###########################################

  modelsvm <- svm(V1~.,dat.tr[, -1062])

  #Predict using SVM regression
  predYsvm = predict(modelsvm, dat.te[, -1062])
  RMSE[i,j] <- calcRMSE(dat.te[,1], predYsvm)
  predictions.te[,j] <- predYsvm
  predictions.tv[,j] <- predict(modelsvm, dat.tv[, -1062])
  j <- j+1

########################################### 

  RMSE[i,j] <-  calcRMSE(dat.te[,1],   rowMeans(predictions.te[, 1:13]))

    
   colnames(predictions.tv) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    colnames(predictions.te) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    
  lmfit<- lm(dat.tv[, 1]~predictions.tv[, 1:13] - 1)
# summary(  lmfit)
  
  pred.stack <- as.matrix( predictions.te[, 1:13]) %*% as.matrix(lmfit$coef)
  # pred.stack <- as.matrix(cbind(rep(1, nrow(x.te)), predictions.te[, 1:13])) %*% as.matrix(lmfit$coef)
   lmfit2<- glmnet(predictions.tv,dat.tv[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
# summary(  lmfit)
  
    pred.stack2 <- predict(lmfit2, predictions.te)
    
    rf.fitSG <- ranger(y = dat.tv[, 1], x = cbind( predictions.tv, dat.tv[, -c(1,1062)]), importance = "impurity")

  pred.rf <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))

# summary(  lmfit)
  
    pred.stack3 <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))$prediction
   RMSE[i,j+1] <- calcRMSE(dat.te[,1],  pred.stack)
    RMSE[i,j+2] <- calcRMSE(dat.te[,1],  pred.stack2)
      RMSE[i,j+3] <- calcRMSE(dat.te[,1],  pred.stack3)
    
  Coefs1[i,] <- coef(lmfit)
  Coefs2[i,] <- coef(lmfit2)[2:14]
  # Coefs3[i,] <- plot(rf.fitSG$variable.importance)
}

colnames(RMSE) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm", "ensMA",  "ensLM",  "ensNonNeg", "ensRF")




RMSE2 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:17,"ALGORITHM")
# RMSE1 %>% ggplot(aes(x = ALGORITHM, y = value)) + geom_boxplot()
RMSE2 %>% group_by(ALGORITHM) %>% mutate(meanv = mean(value)) %>% ggplot(aes(x= ALGORITHM, y = value, color = split, group = split))+
  geom_point()+ geom_line() +geom_line(aes(x = ALGORITHM, y = meanv), col = 1)+ theme(legend.position = "none")+
  ylab("RMSE")+
    theme(axis.text.x=element_text(angle = 90, hjust = 0.5, vjust = 0.5))

t2 <- RMSE2 %>% group_by(ALGORITHM) %>% summarise(meanv = mean(value), sdv=sd(value)) %>% arrange(desc(meanv))


```


## Casein_micelle_size

```{r echo = FALSE}
knitr::kable(t2)
```


## Native_pH



```{r fig.cap="Native_pH", echo=FALSE, message=FALSE, warning=FALSE}


dat <- cbind((y[,3]), tr)
dat <- as.data.frame(dat)
dat <- dat %>% mutate(class = as.factor(labl.Ave))
dat <- dat %>% na.omit()
nonz <- nonz3


# ====replace below
RMSE <- matrix(0, N, 17)
predictions.te <- matrix(0,nrow(dat)-260, 13)
predictions.tv <- matrix(0,130, 13)

Coefs1 <- matrix(0,N, 13)
Coefs2 <- matrix(0,N, 13)
Coefs3 <- matrix(0,N, 13)
i <- 1
for (i in 1:N){
  j <- 1
  dat.tr <- dat[ttr[,i],]
  dat.tv <- dat[ttv[,i],]
  dat.te <- dat[-ttt[,i],]


  x <- model.matrix(V1 ~. , data = dat.tr[, -1062])

  y1 <- dat.tr[,1]

  grid <- 10^seq(-3, 3, length = 100)




  lasso.fit <- glmnet(x,y1,alpha=1, lambda = grid) # for
  cv.out <- cv.glmnet(x,y1,alpha=1)


  lasso.fit <- glmnet(x,y1,alpha=1,
                      lambda = cv.out$lambda.min)



  lasso.pred <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.te[, -1062]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], lasso.pred)
  predictions.te[,j] <- lasso.pred
  predictions.tv[,j] <- predict(lasso.fit, newx = model.matrix(V1 ~. , data = dat.tv[, -1062]))
  j <- j + 1


###########################################




  pc <- prcomp(dat.tr[, 2:1061], scale = TRUE)



  lmfit<- lm(dat.tr[, 1]~pc$x[,1:5])

  x.te <- predict(pc, dat.te[, 2:1061])

  pca.pred.poor.te <- as.matrix(cbind(rep(1, nrow(x.te)), x.te[,1:5])) %*% as.matrix(lmfit$coef[1:6])
  
  x.tv <- predict(pc, dat.tv[, 2:1061])

  pca.pred.poor.tv <- as.matrix(cbind(rep(1, nrow(x.tv)), x.tv[,1:5])) %*% as.matrix(lmfit$coef[1:6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j + 1


###########################################
  pls.fit <- plsr(V1 ~., data=dat.tr[, -1062], scale= TRUE, validation = "CV")

  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 3)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
  predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 3)
  j <- j+ 1

  
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 4)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 4)
  j <- j+ 1
###########################################
  pls.pred <- predict(pls.fit, dat.te[,-c(1, 1062)], ncomp = 5)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pls.pred)
  predictions.te[,j] <- pls.pred
    predictions.tv[,j] <- predict(pls.fit, dat.tv[,-c(1, 1062)], ncomp = 5)
  j <- j+ 1

  
  ###########################################
  lmfit <- lm(dat.tr[, 1]~ as.matrix(dat.tr[, nonz+1]) )


  x <- model.matrix(V1 ~. , data = dat.te[, c(1, nonz+1)])
  pca.pred.poor.te <- x %*% as.matrix(lmfit$coef)
 pca.pred.poor.tv <- model.matrix(V1 ~. , data = dat.tv[, c(1, nonz+1)]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

  ###########################################


  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity")

  pred.rf <- predict(rf.fit , data = dat.te[, -1062])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions 
  j <- j+1


###########################################
  rf.fit <- ranger(V1 ~ ., data = dat.tr[, -1062], importance = "impurity",regularization.factor = 0.2, regularization.usedepth=FALSE)
  # plot(rf.fit$variable.importance)
  pred.rf <- predict(rf.fit , data = dat.te)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
    predictions.tv[,j] <- predict(rf.fit, data = dat.tv[, -1062])$predictions
  j <- j+1

###########################################


  bart_machine = bartMachine(as.data.frame(dat.tr[,-c(1,1062)]), dat.tr[,1], verbose = FALSE)
  # summary(bart_machine)
  pred.bart <- predict(bart_machine, as.data.frame(dat.te[,-c(1,1062)]))
  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.bart)
  predictions.te[,j] <- pred.bart
    predictions.tv[,j] <- predict(bart_machine, as.data.frame(dat.tv[,-c(1, 1062)]))
  j <-j+1

###########################################

  Ktrain <- kernelMatrix(kfunc, as.matrix(dat.tr[,2:1061]))
  Ktest <- kernelMatrix(kfunc, as.matrix(dat.te[,2:1061]), as.matrix(dat.tr[,2:1061]))
   Kvalid <- kernelMatrix(kfunc, as.matrix(dat.tv[,2:1061]), as.matrix(dat.tr[,2:1061]))
 
  kpcTrain <- kPCA(Ktrain)


  kpcTest <- predict(kpcTrain, Ktest)
  kpcValid <- predict(kpcTrain, Kvalid)
  # pairs(kpcTest[ , 1 : 6], col = dat.te$class)
  lmfit <- lm(dat.tr[, 1]~  kpcTrain$KPCs[ , 1 : 6] )





  pca.pred.poor.te <- cbind(rep(1,nrow(kpcTest)), kpcTest[ , 1 : 6]) %*% as.matrix(lmfit$coef)

  pca.pred.poor.tv <- cbind(rep(1,nrow(kpcValid)), kpcValid[ , 1 : 6]) %*% as.matrix(lmfit$coef)


  RMSE[i,j] <- calcRMSE(dat.te[,1], pca.pred.poor.te)
  predictions.te[,j] <- pca.pred.poor.te
  predictions.tv[,j] <- pca.pred.poor.tv
  j <- j+ 1

###########################################


  rf.fit <- ranger(y = dat.tr[, 1], x =  kpcTrain$KPCs[ , 1 : 6], importance = "impurity")

  pred.rf <- predict(rf.fit , data = kpcTest[ , 1 : 6])

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = kpcValid[ , 1 : 6])$predictions
  j <- j+1

###########################################

  rf.fit <- ranger(y = dat.tr[, 1], x =  Ktrain, importance = "impurity")

  pred.rf <- predict(rf.fit , data = Ktest)

  RMSE[i,j] <- calcRMSE(dat.te[,1], pred.rf$predictions)
  predictions.te[,j] <- pred.rf$predictions
  predictions.tv[,j] <- predict(rf.fit , data = Kvalid)$predictions
  j <- j+1

###########################################

  modelsvm <- svm(V1~.,dat.tr[, -1062])

  #Predict using SVM regression
  predYsvm = predict(modelsvm, dat.te[, -1062])
  RMSE[i,j] <- calcRMSE(dat.te[,1], predYsvm)
  predictions.te[,j] <- predYsvm
  predictions.tv[,j] <- predict(modelsvm, dat.tv[, -1062])
  j <- j+1

########################################### 

  RMSE[i,j] <-  calcRMSE(dat.te[,1],   rowMeans(predictions.te[, 1:13]))

    
   colnames(predictions.tv) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    colnames(predictions.te) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm")
    
  lmfit<- lm(dat.tv[, 1]~predictions.tv[, 1:13] - 1)
# summary(  lmfit)
  
  pred.stack <- as.matrix( predictions.te[, 1:13]) %*% as.matrix(lmfit$coef)
  # pred.stack <- as.matrix(cbind(rep(1, nrow(x.te)), predictions.te[, 1:13])) %*% as.matrix(lmfit$coef)
   lmfit2<- glmnet(predictions.tv,dat.tv[,1],alpha=1, lambda = 0, lower.limits = 0, intercept = FALSE) 
# summary(  lmfit)
  
    pred.stack2 <- predict(lmfit2, predictions.te)
    
    rf.fitSG <- ranger(y = dat.tv[, 1], x = cbind( predictions.tv, dat.tv[, -c(1,1062)]), importance = "impurity")

  pred.rf <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))

# summary(  lmfit)
  
    pred.stack3 <- predict(rf.fitSG , data = cbind(predictions.te, dat.te[, -c(1,1062)]))$prediction
   RMSE[i,j+1] <- calcRMSE(dat.te[,1],  pred.stack)
    RMSE[i,j+2] <- calcRMSE(dat.te[,1],  pred.stack2)
      RMSE[i,j+3] <- calcRMSE(dat.te[,1],  pred.stack3)
    
  Coefs1[i,] <- coef(lmfit)
  Coefs2[i,] <- coef(lmfit2)[2:14]
  # Coefs3[i,] <- plot(rf.fitSG$variable.importance)
}


colnames(RMSE) <- c("lasso", "lm+pca", "pls3", "pls4","pls5",  "lm+15", "rf", "rf+vs", "bart", "lm+kpca","rf+kpca", "rf+kernel", "svm", "ensMA",  "ensLM",  "ensNonNeg", "ensRF")




RMSE3 <- RMSE  %>%as.data.frame()%>% mutate(split = as.factor(1:N))%>% pivot_longer(1:17,"ALGORITHM")
# RMSE1 %>% ggplot(aes(x = ALGORITHM, y = value)) + geom_boxplot()
RMSE3 %>% group_by(ALGORITHM) %>% mutate(meanv = mean(value)) %>% ggplot(aes(x= ALGORITHM, y = value, color = split, group = split))+
  geom_point()+ geom_line() +geom_line(aes(x = ALGORITHM, y = meanv), col = 1)+ theme(legend.position = "none")+
  ylab("RMSE")+
    theme(axis.text.x=element_text(angle = 90, hjust = 0.5, vjust = 0.5))

t3 <- RMSE3 %>% group_by(ALGORITHM) %>% summarise(meanv = mean(value), sdv=sd(value)) %>% arrange(desc(meanv))


```


## Native_pH
```{r echo = FALSE}
knitr::kable(t3)
```


## Thank you

- code at https://github.com/domijan/vistamilk

- remove outliers?
